---
title: "RR1 data activity"
author: "Aida Parnia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Importing data

```{r}
# adding the needed libraries for R
library(tidyverse) # for data a cleaning
library(srvyr) # to apply weights
library(kableExtra) # for better tables 
library(modelsummary) # for easier regression tables
```
Reading the data into R. For replication change to your own file location.

```{r}
# Reading data
census <- read_csv("/Users/Oliver/Library/CloudStorage/Dropbox/Sociology/CAnD3/CAND3 Data/Census 2016/pumf-98M0001-E-2016-individuals_F1.csv") # Change path address to your own. 
```
## Cleaning data

### Defining the variables and lables 


**Visible minority** detailed variable called VisMin. 
Defining the categories based on the codebook. 

```{r}

census <- census %>% 
  mutate(VisMin_cat = case_when(
    VisMin == 1 ~ "South Asian",
    VisMin == 2 ~ "Chinese",
    VisMin == 3 ~ "Black",
    VisMin == 4 ~ "Filipino",
    VisMin == 5 ~ "Latin American",
    VisMin == 6 ~ "Arab",
    VisMin > 6 & VisMin < 13 ~ "Other",
    VisMin == 13 ~ "White",
    VisMin == 88 ~ NA_character_
  ))

```

Other variables: Sex, generation status, Age groups, Marital status 

```{r}
# sex
census <- census %>% mutate(Sex_cat = case_when(
  Sex == 1 ~ "Female",
  Sex == 2 ~ "Male"
))

# generation status
census <- census %>% mutate(GENSTAT_cat = case_when(
  GENSTAT == 1 ~ "First",
  GENSTAT %in% c(2,3) ~ "Second",
  GENSTAT == 4 ~ "Third+"
))

# age groups, only those above 65 years old
census <- census %>% mutate(AGEGRP_cat = case_when(
  AGEGRP == 17 ~ "65 to 69 years",
  AGEGRP == 18 ~ "70 to 74 years",
  AGEGRP == 19 ~ "75 to 79 years",
  AGEGRP == 20 ~ "80 to 84 years",
  AGEGRP == 21 ~ "85 years and over"
))

# Marital status
census <- census %>% mutate(MarStH_cat = case_when(
  MarStH == 1 ~ "Never married",
  MarStH == 2 ~ "Married",
  MarStH == 3 ~ "Common law",
  MarStH > 3 ~ "Seperated/divorce/widow"
))

```

Income variables of interest is MrkInc, meaning everything but government transfers. 
The missing are coded as numbers, where 8888888 is not applicable, 99999999 is not available.


```{r}
# Market income cleaned
census <- census %>% mutate(MrkInc_c = case_when(
  MrkInc == 88888888 ~ NA_real_,
  MrkInc == 99999999 ~ NA_real_,
  TRUE ~ MrkInc
))


```

## Descriptive statistics

### Table 1: descriptive

Table 1: the description of the sample by key variables and median market incomes. 
```{r}
# defining analytic sample
# selecting variables of interest and limiting to above 65 years olds, excluding missing
census_AS <- census %>% select(Sex_cat, VisMin_cat, GENSTAT_cat, MrkInc_c, WEIGHT, MarStH_cat, AGEGRP_cat) %>%
  na.omit()

descrip_tab <- census_AS %>% mutate(across(c(where(is.character)), ~ as_factor(.))) %>%
  pivot_longer(-c(MrkInc_c,WEIGHT),names_to = "Variable",values_to = "level") %>%
  as_survey_design(ids = 1, weights = WEIGHT) %>%
  group_by(Variable, level) %>% 
  summarize(n=n(),
            prop_w = round(survey_mean(vartype = "ci")*100,1),
            median_MrkInc = survey_median(MrkInc_c,vartype = "se", na.rm = TRUE)) %>%
  mutate(`Weighted Percentage` = paste0(prop_w," (",format(prop_w_low,1),",",prop_w_upp,")"),
         median_MrkInc = sprintf("$ %.0f", median_MrkInc)) %>% 
  arrange(Variable)

kable(descrip_tab %>% ungroup() %>% select(level,n, `Weighted Percentage`, median_MrkInc),
      col.names = c("", "n", "Weighted %", "Median market income"),
  caption = "Table 1 Descriptive statistics",
  booktabs = FALSE, linesep = "", longtable = FALSE
  ) %>%
  kable_styling(latex_options = c("repeat_header",font_size=12)) %>%
  footnote(general = c("Data from census 2016 (PUMF), those 65 years old and above"))  %>%
  pack_rows(index = c("Age group" = 5, "Generation status" = 3, "Marital status" = 4,
                      "Sex" = 2,"Visible minority" = 8), indent = TRUE)
```


### Figure 1: Income by gender and visible minority group

```{r}

census_AS %>% group_by(Sex_cat,VisMin_cat) %>% summarize(mean_se(MrkInc_c)) %>%
  ggplot(aes(x = VisMin_cat, y = y, group = Sex_cat, fill = Sex_cat)) +
  geom_col(position = "dodge") +
  geom_errorbar(position=position_dodge(.9),width=.25, aes(ymax=ymax, ymin=ymin),alpha=0.3) +
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Market based income by gender and visible minority group",
       caption = "Census 2016 (PUMF), Above 65 years old",
       y = "Market income ($)",
       x = "Visible minoirity group",
       fill = "Sex")
  
```


### Regression results 1

The differences in market income by visible minority groups and sex, adjusted for marital status, generation status, and age groups.

```{r}

# set white as the reference group 
census_AS <- census_AS %>% mutate(VisMin_cat = fct_relevel(VisMin_cat,"White"))

inc_model <- lm(MrkInc_c ~ VisMin_cat * Sex_cat + GENSTAT_cat + MarStH_cat + AGEGRP_cat,
                data = census_AS, weights = WEIGHT)

modelsummary(inc_model, statistic = "conf.int")
```

